Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-14 paper/paper.tex
revised the intro slightly.

Inference algorithms execute Anglican programs while trying to
answer probabilistic queries on those programs. This execution is 
significantly different from the one described in the standard 
operational semantics. These algorithms typically run Anglican programs 
multiple times, often hundreds of thousands or even millions of times 
for a single inference task. The algorithms may make random choices 
that do not correspond to any statements in the program, and decide which parts 
of the program code are executed and how often. Some inference algorithms 
re-run the program multiple times partially, from a certain
point on, while reusing random choices made in the previous runs as much as possible. 
A good high-level picture is that each inference algorithm specifies a 
virtual machine that executes Anglican programs according to a 
non-standard (often probabilistic) operational semantics. Supporting a wide range 
of inference algorithms and their unusual operational semantics efficiently is the main reason 
that we developed Anglican as a language rather than as a library.

@@

\subsection*{Contributions} This paper brings the following
major contributions:
\begin{itemize}
    \item Design and implementation of a probabilistic programming language
        Anglican involving tight bilateral integration with a general-purpose
        programming language.
    \item Specification of probabilistic forms in Anglican which facilitates 
        separation of probabilistic programs from inference algorithms.
    \item Stateless representation of random processes, seamlessly integrated
        into pure functional computation.
    \item Techniques for efficient and compact implementation of inference
        algorithms, such as representation of inference results as a lazy
        sequence of samples and a novel scheme for addressing of checkpoints.
\end{itemize}

>>>>

Inference algorithms execute Anglican programs while trying to
answer probabilistic queries on those programs. This execution is
significantly different from the one described in the standard 
operational semantics. These algorithms typically run Anglican programs 
multiple times, often hundreds of thousands or even millions of times 
for a single inference task. The algorithms may make random choices 
that do not correspond to any statements in the program, and decide which parts 
of the program code are executed and how often. Some inference algorithms 
re-run the program multiple times partially, from a certain
point on, while reusing random choices made in the previous runs as much as possible. A good high-level picture is that each inference algorithm specifies a 
virtual machine that executes Anglican programs according to a 
non-standard (usually probabilistic) operational semantics. 
Supporting a wide range of inference algorithms and their unusual semantics 
the main reason that we developed Anglican as a language rather 
than as a library.

@@

\subsection*{Contributions} This paper brings the following
major contributions:
\begin{itemize}
    \item Design and implementation of a probabilistic programming language
        Anglican involving tight bilateral integration with a general-purpose
        programming language.
    \item Techniques for efficient and compact implementation of inference
        algorithms, such as representation of inference results as a lazy
        sequence of samples and a novel scheme for addressing of checkpoints.
    \item Anglican's novel representation of random processes 
            from statistics and machine learning, such as Beta-Binomial 
            process. The representation is
            \emph{stateless} and seamlessly integrates into pure functional 
            computation. This contrasts with typical \emph{stateful}
            implementations of these processes in other probabilistic 
            programming languages.
            These constructs allowrepresents these p
%    \item Specification of probabilistic forms in Anglican which facilitates 
%        separation of probabilistic programs from inference algorithms.
\end{itemize}
